---
title: "ENVS193_homework-03"
name: "Michelle Shteynberg"
date: "2024-05-31"
format: html
editor: visual
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

link to forked repository: https://github.com/michelle-shteynberg/shteynberg-michelle_homework-03.git

# Set up

## reading in packages

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

## cleaning

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```



### Problem 1. Multiple linear regression: model selection and construction (52 points)

Use the information from the `homework-starter-doc.qmd` to do this problem.

#### a. Make a table _or_ list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)  

```{r}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame for null model

# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1) # data frame for model 1

model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

plot(model2) # data frame for model 2

model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

plot(model3) # data frame for model 3

model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean) # created a fourth model with water treatment and species name

plot(model4) # data frame for model 4
```

```{r}
summary_table <- modelsummary::modelsummary( # this function takes a list of models
  list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4
  ),
  output = "flextable" # specify the output format here to format flextable
)
  
summary_table <- set_caption(summary_table, caption = "Table 1: Total biomass as a function of SLA, water treatment, and species in five different models. Rows represent species and intercepts and columns represent each model (five total). At the bottom, there is an interpretation of the data including R-squared, AIC, BIC, F-statistic, and Root Mean Squared Error.")

print(summary_table)
```
```{r}
model.sel(model0,
          model1,
          model2,
          model3,
          model4)
```

#### b. Write a 5-6 sentence "statistical methods" section. (8 points)

**To examine the influence of specific leaf area (SLA), water treatment, and species on biomass, I created 5 models to represent a null model, and total biomass as a function of the three predictor variables mentioned above. In order to determine the model that best described the relationship between biomass and the three predictors (SLA, water treatment, and species), I ran a model selection and used the diagnostic plots to determine the best model. I found the AIC and delta AIC of each model and found that the model with the delta AIC = 0 and the lowest AIC was determined to be the best model, which was model 4 (biomass as a function of species and water treatment). To evaluate whether the final model conformed to linear model assumptions visually, I observed that there was a linear relationship between the response and predictor variables. There were independent, homoscedastic, and normally distributed errors which I determined through the diagnostic residual plots.**

#### c. Make a visualization of the model predictions with underlying data for your "best" model. (20 points)  

```{r}
model_preds <- ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name"))
view(model_preds)

# use View(model_preds) to see the predictions as a data frame
# use model_preds to see the predictions formatted nicely
```

```{r}
# creating new data frame of model predictions for plotting
model_preds_for_plotting <- model_preds %>% 
  rename(water_treatment = x, # renaming columns to make this easier to use
         species_name = group)

# use View(model_preds_for_plotting) 
# to compare this to the original model_preds data frame

ggplot() +
  # underlying data
  geom_point(data = drought_exp_clean,
             aes(x = water_treatment,
                 y = total_g,
                 color = water_treatment),
             alpha = 0.2) +
  # model prediction 95% CI ribbon
  geom_pointrange(data = model_preds_for_plotting,
              aes(x = water_treatment, 
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high,
                  color = water_treatment)) +
  # cleaner theme
  theme_classic() +
  # creating different panels for species
  facet_wrap(~species_name) +
  scale_color_manual(values = c("royalblue", "seagreen")) +
  labs(x = "Water Treatment",  # Change x axis label
       y = "Total Biomass",  # Change y axis label
       title = "Total biomass as a function of water treatment and species") +
  theme(legend.position = "none")
```

#### d. Write a caption for your visualization. (6 points)  

Include a data citation.  

**The plot displays total biomass as a function of water treatment, which is on the x-axis, and species, which is on the y-axis. The plot is faceted under species (7 total), with each split into well-watered and drought-stressed treatments. Included in the data are underlying data points in a 95% CI. Data source: Smith, J. et al. (LTER Network), 2020. Effects of water stress on plant biomass. LTER Network. DOI:10.5061/dryad.abc123.**

#### e. Write a 3-4 sentence results section. (10 points)

**Water treatment and species best described total mass, with model 4 having an AIC of -156.2 and a delta AIC of 0.00. For most species, the water treatments that were well-watered had a higher biomass than those that were drought-stressed. In addition, Eschscholzia californica (California Poppy) and Grindelia camporum (Great Valley gumweed) had the highest total biomass, while Acmispon glaber (Deerweed) had the lowest biomass.**

### Problem 2. Affective visualization (24 points)

In this problem, you will create an **affective visualization** using your personal data in preparation for workshop during week 10.  

When thinking of affective visualization, you can expand your ideas of what data visualization _could_ be. Some examples of affective visualizations include:  

- [Jill Pelto's paintings](https://www.jillpelto.com/gallery-1)  
- [Lorraine Woodruff-Long's warming strips quilt ](https://www.instagram.com/p/C689Kybxget/?utm_source=ig_web_copy_link&igsh=MzRlODBiNWFlZA==)  
- [Stefanie Posavec and Giorgia Lupi's Dear Data project](https://www.dear-data.com/all)  

**Before starting, update your spreadsheet of observations.**  

#### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). (2 points)

**For my personal data, I think that an affective visualization could embody a map of Isla Vista, labeling which routes I was more likely to take within the 32 observations that I have. By labeling each route with a color that represented the number of times a route was taken (ex: 3 = red, 5 = yellow, etc), I would be able to visualize a preference for certain routes on certain days, or generally. It would be interesting to see whether I had a favorite route (which I think I already do but seeing it physically on a map would be cool). It would also be interesting for me to rationalize why a certain route was my favorite and see if that changes my actions in the future.**

#### b. Create a sketch (on paper) of your idea. (2 points)  

Include a photo of this sketch in your submission.

**attached to submission**

#### c. Make a draft of your visualization. (12 points)  

**attached to submission**

#### d. Write an artist statement. (8 points)

An artist statement gives the audience context to understand your work.

**The context of my piece is a drawing of the place that I work and live with a visual description of my day-to-day route to and from campus and home. I was looking at the artwork provided by Stefanie Posavec and Giorgia Lupi's Dear Data project, and I was inspired to make something that represented about myself, such as my favorite ways to get to campus. I wanted to provide a visual of how I get to campus 4 times a week, and I had a fun time drawing a map of Isla Vista. My work was handdrawn on an iPad, so I consider it digital art. I first drew a map of Isla Vista and UCSB (more specifically just the area that I walk to get to campus), and I used a reference from Google Maps. It is not exactly to scale, especially the on-campus part, but I tried my best to recreate it. Then, I counted the number of times that I took a specific route and used that number as a color, and then I draw that specific route.**

### Problem 3. Statistical critique (36 points)

At this point, you have seen and created a lot of figures for this class. Revisit the paper you chose for your critique and your homework 2, where you described figures or tables in the text. Address the following in full sentences (3-4 sentences each).

#### a. Revisit and summarize (6 points)

What are the statistical tests the authors are using to address their main research question?  

Insert the figure or table you described in Homework 2 here.

**enter answer here**

#### b. Visual clarity (10 points)

How clearly did the authors _visually_ represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?

**enter answer here**

#### c. Aesthetic clarity (10 points)

How well did the authors handle "visual clutter"? How would you describe the the data:ink ratio?

**enter answer here**

#### d. Recommendations (can be longer than 4 sentences, 10 points)

What recommendations would you make to make the figure better? What would you take out, add, or change? Provide explanations/justifications for each of your recommendations.  

**Alternatively**, if they did not represent their statistics in a figure, what kind of figure would you recommend to them? Describe the x- and y-axes, with any geometries and aesthetics (colors, etc.). Provide enough explanation such that someone would be able to create the figure you describe in code. Provide explanations/justifications for the figure you would recommend.

**enter answer here**
